Efficient evaluation of threshold queries of derived fields in a numerical simulation database

### 2015-01-15 paper.tex
Minor fixes.

Cache hits reduce the running time of threshold queries by over an order of magnitude as shown in Figure \ref{fig:cache_overhead} and
Table \ref{tab:cache_effectiveness}. This is because we do not have to compute the requested derived
field from the raw data, which eliminates the associated I/O. Only the cache entries need to be looked-up, which is substantially less data
than the raw vector or scalar field data. For the queries with large result sets it is actually the network time taken to transfer the results to the user that 
dominates the overall execution as opposed to the I/O or computation time as we show later.
Cache hits are evaluated by first warming up the cache by submitting the same set of threshold queries of the vorticity field as before. We then submit
several more unrelated queries with different time-steps and threshold values in order to pollute the cache. Finally, we issue the original set of queries and 
measure their running times. Let us focus on the query
with low threshold, which returns $\sim$900,000 points. Given that valid threshold values are limited to those that result in no more than 1,000,000 points
it is likely that all subsequent queries to this time-step will result in a cache hit as their threshold is likely to be equal or higher than the cached one.

@@

\subsection{Evaluation of Additional Fields}
The data-parallel evaluation of threshold queries shows stable execution time for different derived fields in addition to the different threshold levels and
time-steps queried. The execution times depend on the complexity of the computation needed to evaluate the particular derived field requested. Figures 
\ref{fig:vorticity_breakdown}, \ref{fig:q_breakdown}, \ref{fig:magnetic_breakdown} show a breakdown of the execution time of threshold queries of different
derived fields, which are evaluated from the raw data and a cold cache using 4 processes per node on a 4 node cluster. Almost the entire time is spent
performing the I/O and computation associated with the derived field requested.

@@

The processing of top-$k$ queries has been studied extensively in the context of distributed and relational database systems. A survey of different techniques
in the case of centralized processing is given in \cite{Ilyas}. In the case of distributed processing different approaches focus on horizontally 
\cite{Balke, Vlachou} or vertically \cite{Cao, Chaudhuri, Guntzer, Marian, Michel} distributed data. None of these approaches deals with array data stored in
a relational database system. Zhao et al. propose an algorithm for the processing of top-$k$ queries in large-scale distributed environments called BRANCA
\cite{Zhao}. They build on the idea of semantic caching \cite{Ren} and make use of branch caches, which store results of previous top-$k$ queries with respect 
to the data stored on each server. The caching mechanism that we use is similar in that regard, but the queries that are evaluated in our system operate on
derived fields, which are computed at each location by accessing data from a surrounding region. The queries described in \cite{Zhao} operate over the 
attributes of individual records only using simple linear score functions.

@@

Sampling approaches \cite{Sidirourgos, Nirkhiwale} offer an alternative to the on-demand computation of derived fields and the evaluation of threshold queries
on them. The goal of both techniques is to not return large data volumes, but focus on the most intense events and interesting regions in the dataset. 
The computation of the derived fields is carried out on the nodes of the database cluster and takes a look at the dataset as a whole, while the user obtains only
a small subset of the data, where the derived field in question is above the prescribed threshold. Sampling approaches can potentially omit some locations and
while useful for generating initial impressions may not be suitable if the exact locations where a field is at its highest values are desired.

>>>>

Cache hits reduce the running time of threshold queries by over an order of magnitude as shown in Figure \ref{fig:cache_overhead} and
Table \ref{tab:cache_effectiveness}. This is because we do not have to compute the requested derived
field from the raw data, which eliminates the associated I/O. Only the cache entries need to be looked-up, which is substantially less data
than the raw vector or scalar field data. For the queries with large result sets it is actually the network time taken to transfer the results to the user that 
dominates the overall execution as opposed to the I/O or computation time as we show later.
Cache hits are evaluated by first warming up the cache by submitting the same set of threshold queries of the vorticity field as before. We then submit
several more unrelated queries with different time-steps and threshold values in order to pollute the cache. Finally, we issue the original set of queries and 
measure their running times. Let us focus on the query
with low threshold, which returns $\sim$900,000 points. Given that valid threshold values are limited to those that result in no more than 1,000,000 points
it is likely that all subsequent queries to this time-step will result in a cache hit as their threshold is likely to be equal or higher than the cached one.
Currently we observe fairly high cache-hit ratios as the workload is very structured and queries tend to examine the same regions in space and time.

@@

\subsection{Evaluation of Additional Fields}
The data-parallel evaluation of threshold queries shows stable execution time for different derived fields in addition to the different threshold levels and
time-steps queried. The execution times depend on the complexity of the computation needed to evaluate the particular derived field requested. Figures 
\ref{fig:vorticity_breakdown}, \ref{fig:q_breakdown} and \ref{fig:magnetic_breakdown} show a breakdown of the execution time of threshold queries of different
derived fields, which are evaluated from the raw data and a cold cache using 4 processes per node on a 4 node cluster. Almost the entire time is spent
performing the I/O and computation associated with the derived field requested.

@@

The processing of top-$k$ queries has been studied extensively in the context of distributed and relational database systems. A survey of different techniques
in the case of centralized processing is given in \cite{Ilyas}. In the case of distributed processing different approaches focus on horizontally 
\cite{Balke, Vlachou} or vertically \cite{Cao, Chaudhuri, Guntzer, Marian, Michel} distributed data. None of these approaches deal with array data stored in
a relational database system. Zhao et al. propose an algorithm for the processing of top-$k$ queries in large-scale distributed environments called BRANCA
\cite{Zhao}. They build on the idea of semantic caching \cite{Ren} and make use of branch caches, which store results of previous top-$k$ queries with respect 
to the data stored on each server. The caching mechanism that we use is similar in that regard, but the queries that are evaluated in our system operate on
derived fields, which are computed at each location by accessing data from a surrounding region. The queries described in \cite{Zhao} operate over the 
attributes of individual records only using simple linear score functions.

@@

Sampling approaches \cite{Sidirourgos, Nirkhiwale} offer an alternative to the on-demand computation of derived fields and the evaluation of threshold queries
on them. The goal of both techniques is to not return large data volumes, but focus on the most intense events and interesting regions in the dataset. 
The computation of derived fields is carried out on the nodes of the database cluster and takes a look at the dataset as a whole, while the user obtains only
a small subset of the data, where the derived field in question is above the prescribed threshold. Sampling approaches can potentially omit some locations and
while useful for generating initial impressions may not be suitable if the exact locations where a field is at its highest values are desired.
